---
title: "Metalimnion depth calculations RMD"
author: "Thomas Van Zuiden"
date: "June 27, 2017"
output: html_document
---


### Load packages
```{r warning=FALSE, message=FALSE}
library(rLakeAnalyzer)
library(dplyr)
```

### Load data
```{r echo=FALSE, results='hide'}
setwd("C:\\Users\\adminuser\\Desktop\\EVERYTHING THOMAS\\MOECC Alagae project\\Lake Physics")
```
```{r results='hide'}
dat <- read.csv("tempDO_profiles.csv")
names(dat)
dat <- dat[, 1:8] # left out the dissolved oxygen column because it had a lot of NAs
dat <- na.omit(dat) # omit rows with missing temperature and depth values
```

### Using the "meta.depths" function in rLakeAnalyzer to create a simple function with set parameters
```{r}
meta <- function (x) {
  meta.depths(x$TEMPERATURE, x$DEPTH, slope=0.1, seasonal=FALSE)
}

test <- dat[1:21,] ### These rows represent 1 sampling period
meta(test)
### Top of metalimnion = 7.017519m
### Bottom of metalimnion = 7.750571m
```

 
### Note 1: I create these new variables for individual lakes because some had data inputted incorrectly and it was easier to fix when working lake by lake

### Note 2: The reason I calculated metalimnion depths for August was because it is the only month each year that has data for all years for each lake between 1981-2015. Other months range from missing one to several years of data in each lake

### Note 3: As I ran through each lake, I found a data entry mistake in Harp Lake. Steps on how to fix it are included

### Note 4: Heney Lake is extremely shallow, and consequently, produces a lot of NAs; presumably because the "meta" function cannot find the metalimnion


## BLUE CHALK LAKE
```{r}
BC <- dat[dat$LAKE_NAME=="Blue_Chalk", 1:8]
BC8 <- filter(BC, MONTH == 8)
df1 <- data.frame() # Creating an empty data frame

sampling.BC <- unique(BC8$SAMPLE_DATE)
```

### Create a loop that uses the "meta" function to calculate the depth of the top and bottom of the metalimnion for each unique sampling date
```{r}
for (i in sampling.BC) {
  m <- meta(BC8[BC8$SAMPLE_DATE==i, 4:5])
  outputs <- data.frame(sample_date=i, m)
  df1 <- rbind(df1,outputs) # Putting outputs in our empty data frame created above
}
```

#### If you look at "df1" after running the loop you'll see that top and bottom metalimnion depths are stacked on top of each other in a single column.The following code separates the top and bottom metalimnion depths and puts them in separate columns
```{r}
df.new1 = df1[seq(1, nrow(df1), 2), ]
df.new2 = df1[seq(2, nrow(df1), 2), ]

df.BC <- data.frame(df.new1, df.new2)
df.BC <- df.BC[, -c(3), drop=FALSE]
colnames(df.BC)[2] <- "Top_metalimnion"
colnames(df.BC)[3] <- "Bottom_metalimnion"

df.BC$Lake_Name <- c("Blue Chalk")
df.BC<-df.BC[, c(4,1:3)]
head(df.BC)
```

## CHUB LAKE
```{r}
CB <- dat[dat$LAKE_NAME=="Chub", 1:8]
CB8 <- filter(CB, MONTH == 8)
df.CB8 <- data.frame()

sampling.CB <- unique(CB8$SAMPLE_DATE)
for (i in sampling.CB) {
  m1 <- meta(CB8[CB8$SAMPLE_DATE==i, 4:5])
  outputs <- data.frame(sample_date=i, m1)
  df.CB8 <- rbind(df.CB8,outputs)
}

df.CB8.new1 = df.CB8[seq(1, nrow(df.CB8), 2), ]
df.CB8.new2 = df.CB8[seq(2, nrow(df.CB8), 2), ]

df.CB <- data.frame(df.CB8.new1, df.CB8.new2)
df.CB <- df.CB[, -c(3), drop=FALSE]
colnames(df.CB)[2] <- "Top_metalimnion"
colnames(df.CB)[3] <- "Bottom_metalimnion"

df.CB$Lake_Name <- c("Chub")
df.CB<-df.CB[, c(4,1:3)]
head(df.CB)
```

## CROSSON LAKE
```{r}
CN <- dat[dat$LAKE_NAME=="Crosson", 1:8]
CN8 <- filter(CN, MONTH == 8)
df.CN8 <- data.frame()

sampling.CN <- unique(CN8$SAMPLE_DATE)
for (i in sampling.CN) {
  m1 <- meta(CN8[CN8$SAMPLE_DATE==i, 4:5])
  outputs <- data.frame(sample_date=i, m1)
  df.CN8 <- rbind(df.CN8,outputs)
}

df.CN8.new1 = df.CN8[seq(1, nrow(df.CN8), 2), ]
df.CN8.new2 = df.CN8[seq(2, nrow(df.CN8), 2), ]

df.CN <- data.frame(df.CN8.new1, df.CN8.new2)
df.CN <- df.CN[, -c(3), drop=FALSE]
colnames(df.CN)[2] <- "Top_metalimnion"
colnames(df.CN)[3] <- "Bottom_metalimnion"

df.CN$Lake_Name <- c("Crosson")
df.CN<-df.CN[, c(4,1:3)]
head(df.CN)
```

## DICKIE LAKE
```{r}
DE <- dat[dat$LAKE_NAME=="Dickie", 1:8]
DE8 <- filter(DE, MONTH == 8)
df.DE8 <- data.frame()

sampling.DE <- unique(DE8$SAMPLE_DATE)
for (i in sampling.DE) {
  m1 <- meta(DE8[DE8$SAMPLE_DATE==i, 4:5])
  outputs <- data.frame(sample_date=i, m1)
  df.DE8 <- rbind(df.DE8,outputs)
}

df.DE8.new1 = df.DE8[seq(1, nrow(df.DE8), 2), ]
df.DE8.new2 = df.DE8[seq(2, nrow(df.DE8), 2), ]

df.DE <- data.frame(df.DE8.new1, df.DE8.new2)
df.DE <- df.DE[, -c(3), drop=FALSE]
colnames(df.DE)[2] <- "Top_metalimnion"
colnames(df.DE)[3] <- "Bottom_metalimnion"

df.DE$Lake_Name <- c("Dickie")
df.DE<-df.DE[, c(4,1:3)]
head(df.DE)
```

## HARP LAKE
```{r}
HP <- dat[dat$LAKE_NAME=="Harp", 1:8]
HP8 <- filter(HP, MONTH == 8)
df.HP8 <- data.frame()
```
### Stop here... this lake had data inputted incorrectly somewhere. I added the column "diff" which is each depth value minus the previous depth value
```{r results="hide"}
HP8$diff <- ave(HP8$DEPTH, HP8$LAKE_NAME, FUN=function(x) c(0, diff(x))) # create "diff"
HP8_error <- HP8[order(HP8$diff),] # Sort by "diff"
HP8_error[1:75,] # Found a 0 in row 945, which means it was a duplicate
HP8[942:947, ] # row 944 and 945 are duplicates
```
### The 27m "DEPTH" profile was duplicated in row 945. This will cause the "meta" function to error out


### Recreate the original dataframe with the duplicate row and the "diff" column removed
```{r}
HP8 <- HP8[-c(945),]
HP8 <- subset(HP8, select = -c(diff))
HP8[942:947, ] # making sure it's gone
```

### The looped "meta" function should work now that the duplicate depth is removed
```{r}
sampling.HP <- unique(HP8$SAMPLE_DATE)
for (i in sampling.HP) {
  m1 <- meta(HP8[HP8$SAMPLE_DATE==i, 4:5])
  outputs <- data.frame(sample_date=i, m1)
  df.HP8 <- rbind(df.HP8,outputs)
}

df.HP8.new1 = df.HP8[seq(1, nrow(df.HP8), 2), ]
df.HP8.new2 = df.HP8[seq(2, nrow(df.HP8), 2), ]

df.HP <- data.frame(df.HP8.new1, df.HP8.new2)
df.HP <- df.HP[, -c(3), drop=FALSE]
colnames(df.HP)[2] <- "Top_metalimnion"
colnames(df.HP)[3] <- "Bottom_metalimnion"

df.HP$Lake_Name <- c("Harp")
df.HP<-df.HP[, c(4,1:3)]
head(df.HP)
```

## HENEY LAKE
```{r}
HY <- dat[dat$LAKE_NAME=="Heney", 1:8]
HY8 <- filter(HY, MONTH == 8)
df.HY8 <- data.frame()

sampling.HY <- unique(HY8$SAMPLE_DATE)
for (i in sampling.HY) {
  m1 <- meta(HY8[HY8$SAMPLE_DATE==i, 4:5])
  outputs <- data.frame(sample_date=i, m1)
  df.HY8 <- rbind(df.HY8,outputs)
}

df.HY8.new1 = df.HY8[seq(1, nrow(df.HY8), 2), ]
df.HY8.new2 = df.HY8[seq(2, nrow(df.HY8), 2), ]

df.HY <- data.frame(df.HY8.new1, df.HY8.new2)
df.HY <- df.HY[, -c(3), drop=FALSE]
colnames(df.HY)[2] <- "Top_metalimnion"
colnames(df.HY)[3] <- "Bottom_metalimnion"

df.HY$Lake_Name <- c("Heney")
df.HY<-df.HY[, c(4,1:3)]
head(df.HY)
```

## PLASTIC LAKE
```{r}
PC <- dat[dat$LAKE_NAME=="Plastic", 1:8]
PC8 <- filter(PC, MONTH == 8)
df.PC8 <- data.frame()

sampling.PC <- unique(PC8$SAMPLE_DATE)
for (i in sampling.PC) {
  m1 <- meta(PC8[PC8$SAMPLE_DATE==i, 4:5])
  outputs <- data.frame(sample_date=i, m1)
  df.PC8 <- rbind(df.PC8,outputs)
}

df.PC8.new1 = df.PC8[seq(1, nrow(df.PC8), 2), ]
df.PC8.new2 = df.PC8[seq(2, nrow(df.PC8), 2), ]

df.PC <- data.frame(df.PC8.new1, df.PC8.new2)
df.PC <- df.PC[, -c(3), drop=FALSE]
colnames(df.PC)[2] <- "Top_metalimnion"
colnames(df.PC)[3] <- "Bottom_metalimnion"

df.PC$Lake_Name <- c("Plastic")
df.PC<-df.PC[, c(4,1:3)]
head(df.PC)
```

## RED CHALK LAKE EAST
```{r}
RE <- dat[dat$LAKE_NAME=="Red_Chalk_E", 1:8]
RE8 <- filter(RE, MONTH == 8)
df.RE8 <- data.frame()

sampling.RE <- unique(RE8$SAMPLE_DATE)
for (i in sampling.RE) {
  m1 <- meta(RE8[RE8$SAMPLE_DATE==i, 4:5])
  outputs <- data.frame(sample_date=i, m1)
  df.RE8 <- rbind(df.RE8,outputs)
}

df.RE8.new1 = df.RE8[seq(1, nrow(df.RE8), 2), ]
df.RE8.new2 = df.RE8[seq(2, nrow(df.RE8), 2), ]

df.RE <- data.frame(df.RE8.new1, df.RE8.new2)
df.RE <- df.RE[, -c(3), drop=FALSE]
colnames(df.RE)[2] <- "Top_metalimnion"
colnames(df.RE)[3] <- "Bottom_metalimnion"

df.RE$Lake_Name <- c("Red Chalk East")
df.RE<-df.RE[, c(4,1:3)]
head(df.RE)
```

## RED CHALK LAKE MAIN
```{r}
RM <- dat[dat$LAKE_NAME=="Red_Chalk_M", 1:8]
RM8 <- filter(RM, MONTH == 8)
df.RM8 <- data.frame()

sampling.RM <- unique(RM8$SAMPLE_DATE)
for (i in sampling.RM) {
  m1 <- meta(RM8[RM8$SAMPLE_DATE==i, 4:5])
  outputs <- data.frame(sample_date=i, m1)
  df.RM8 <- rbind(df.RM8,outputs)
}

df.RM8.new1 = df.RM8[seq(1, nrow(df.RM8), 2), ]
df.RM8.new2 = df.RM8[seq(2, nrow(df.RM8), 2), ]

df.RM <- data.frame(df.RM8.new1, df.RM8.new2)
df.RM <- df.RM[, -c(3), drop=FALSE]
colnames(df.RM)[2] <- "Top_metalimnion"
colnames(df.RM)[3] <- "Bottom_metalimnion"

df.RM$Lake_Name <- c("Red Chalk Main")
df.RM<-df.RM[, c(4,1:3)]
head(df.RM)
```

### Export data
```{r}
all.lakes <- rbind(df.BC,df.CB,df.CN,df.DE,df.HP,df.HY,df.PC,df.RE,df.RM) # stack rows
rownames(all.lakes) <- c() # Just resetting the column of row names
write.csv(all.lakes, "Aug_meta_depths.csv") # Write outputs to a CSV
```